---
title: "S101 HMS Sandbar Shark Rec Discards EDA"
author: "Samantha Binion-Rock"
date: "`r format(Sys.Date(), '%m-%d-%Y')`"
html_notebook:
page-layout: full
toc: true
toc-location: left
editor: visual
---

## 

# Discard Lengths EDA

```{r}
#| output: false
#| echo: false
#| warning: false
#| label: 'load packages'


librarian::shelf(tidyverse, ggplot2, dplyr, gt, openxlsx, readxl, haven, doBy, xtable, reshape2, ExPanDaR, png, grid, gridExtra, ggpubr, withr, flextable, officer, janitor)



```

```{r}
#| output: false
#| echo: false
#| warning: false
#| label: 'load data'

##* *Define directory*
dir <- getwd()

##* *Prevents R from converting any long numbers (ID codes) to scientific notation*
options(scipen=999)


##* *Import data*
lengths <- readxl::read_xlsx(paste0(dir,'/DiscardLenComps/SBS_rec_dsizAPAIS_0424_20260609.xlsx'), sheet = 'sbs', guess_max = 500000)




##* *Order states geographically*
unique(lengths$State_Landed)

lengths$State_Landed <- factor(lengths$State_Landed,
                               levels=c('NJ', 'DE', 'MD', 'VA',
                                        'NC', 'SC', 'GA', 'FL'))




##* *Standardize disposition*
unique(lengths$Disposition)


##* *Remove spaces and commas from Disposition so I can recode*
#lengths$Disposition <- gsub(",","",as.character(lengths$Disposition))

#lengths$Disposition <- gsub(" ","_",as.character(lengths$Disposition))



#lengths$Disposition <- recode(lengths$Disposition,
#                           D                        = 'Discard',
#                           H                        = 'Harvest',
#                           Released_alive_not_legal = 'Discard',
#                           Kept_to_eat              = 'Harvest',
#                           Released_dead            = 'Harvest',
#                           Released_alive_legal     = 'Discard',
#                           Preyed_upon              = 'Discard')




##* *Standardize Fishing Mode*


#unique(lengths$Fishing_Mode)


#lengths$Fishing_Mode <- recode(lengths$Fishing_Mode,
#                           Charter  = 'CB',
#                           Headboat = 'HB',
#                           C        = 'CB')




#lengths$trip_type <- factor(lengths$trip_type,
#                            levels=c('H', 'Q', 'F', 'M'))


lengths$Final_Length_mm <- lengths$Observed_FL_mm



##* *Drop samples with no lengths*
lengths <- lengths %>%
  mutate(Length_Flag = case_when(
    Final_Length_mm > 1 ~ 0,
    is.na(Final_Length_mm) ~ 1))


check <- lengths %>%
  filter(Length_Flag == 1) ## 0 samples


lengths <- lengths %>%
  filter(Length_Flag == 0)







##* *There is an error in the data from MS*
##* *Max total length is in the length field, not to replace with FL values*

##* *Make data set with only MS*
#ms <- lengths %>%
#  filter(State_Landed == 'MS')

#unique(lengths$State_Landed)


##* *Remove MS from lengths*
#lengths <- lengths %>%
#  filter(State_Landed %in% c('AL', 'FL'))


##* *Replace max length with FL for MS*
#ms <- subset(ms, select = -Final_Length_mm)

#ms$Final_Length_mm <- ms$FORK_LENGTH


##* *Add ms back to lengths*
#lengths <- rbind(lengths, ms)



##* *Define length type*
length.type <- 'FL (mm)'
```

```{r}
#| output: false
#| echo: false
#| warning: false
#| eval: true



##* *Define size bins - cm*
min(lengths$Final_Length_mm, na.rm=TRUE)
max(lengths$Final_Length_mm, na.rm=TRUE)


bin_size <- 5

min_size <- 5  

len_plus <- 125

up_bin_limit <- 130 ##defining as len_plus + bin_size (e.g. 121 + 1 cm)



##* *Define 1 cm size bins*
##* *For SA, need to use rnd function to ensure proper rounding (0.5 gets rounded up)*
rnd = function(x) trunc(x+0.5)

#rec.lengths$Final_Length_cm <- rnd(x=rec.lengths$Final_Length_mm/10)


##* *For GOM, use trunc function to label the floor*

lengths$Final_Length_cm <- trunc(lengths$Final_Length_mm/10)




##* *Creating proper size bins based on assessment parameters*
##* *Labeling floor of bin for GOM and middle for SA*
lengths <- lengths %>%
  mutate(bin_range = cut(Final_Length_cm,
                         breaks = seq(min_size, up_bin_limit, 
                                      by=bin_size),
                         include.lowest = T,
                         right = F),
         bin       = cut(Final_Length_cm,
                         breaks = seq(min_size, up_bin_limit, 
                                      by=bin_size),
                         labels = seq(min_size, len_plus, 
                                      by=bin_size),
                         include.lowest = T,
                         right = F))





min(lengths$Final_Length_mm)
max(lengths$Final_Length_mm)




##* *Assign plus bins*
lengths.na <- lengths %>%
  filter(is.na(bin))   ##all lengths assigned

#lengths.na <- lengths.na[,1:22]

#lengths.na <- lengths %>%
#  mutate(bin = case_when(
#         Final_Length_cm < 21 ~ 21,
#         Final_Length_cm > 99 ~ 99,
#         .default = NA))



##* *Drop samples with unassigned bins from lengths*
#lengths <- lengths %>%
#  filter(!is.na(bin))


##* *Combine lengths and lengths.na*
#lengths <- rbind(lengths, lengths.na)





##* *Convert Fork TL in mm to Fork TL in inches*
lengths$FL_inch <- lengths$Final_Length_mm * 0.0393701


##* *Create 1 inch length bin*
lengths$bin_1in <- rnd(x=lengths$FL_inch)




##* *Assign potential subweighting region of MS, AL, NWFL vs SWFL*
#lengths <- lengths %>%
#  mutate(Sub_Reg = case_when(
#    State_Landed %in% c('MS', 'AL') ~ 'North',
#    hb_reg == 'NWFL' ~ 'North',
#    hb_reg == 'SWFL' ~ 'South'))



```

## Sample Sizes by Data Provider

```{r}
#| echo: false

#unique(lengths$State_Landed)



paste0("Table: Table: Number of discarded fish lengths from each data provider.")
lengths %>%
  filter(Disposition == 'Discard') %>%
  tabyl(Year, Data_Provider) %>%
  adorn_totals(c("row", "col"))


```

```{r}
#| echo: false

#unique(lengths$State_Landed)



paste0("Table: Table: Number of harvested fish lengths from each data provider.")
lengths %>%
  filter(Disposition == 'Harvest') %>%
  tabyl(Year, Data_Provider) %>%
  adorn_totals(c("row", "col"))


```

## Sample Sizes by State - Harvest

```{r}
#| echo: false

#unique(lengths$State_Landed)



paste0("Table: Table: Number of harvested fish lengths from each state.")
lengths %>%
  filter(Disposition == 'Harvest') %>%
  tabyl(Year, State_Landed) %>%
  adorn_totals(c("row", "col"))

```

## Sample Sizes by State - Discards

```{r}
#| echo: false

#unique(lengths$State_Landed)



paste0("Table: Number of discarded fish lengths from each state.")
lengths %>%
  filter(Disposition == 'Discard') %>%
  tabyl(Year, State_Landed) %>%
  adorn_totals(c("row", "col"))

```

## Sample Sizes by Disposition

```{r}
#| echo: false




paste0("Table: Number of harvested and discarded fish lengths.")
lengths %>%
  tabyl(Year, Disposition) %>%
  adorn_totals(c("row", "col"))

```

## Sample Sizes - Discards

```{r}
#| echo: false




paste0("Table: Number of discarded fish lengths by trip type.")
lengths %>%
  filter(Disposition == 'Discard') %>%
  tabyl(Year, Fishing_Mode) %>%
  adorn_totals(c("row", "col"))

```

## Sample Sizes by Trip Type

```{r}
#| echo: false
#| eval: false

#unique(lengths$State_Landed)



paste0("Table: Number of discarded fish lengths from each trip type.")
lengths %>%
  filter(Disposition == 'Discard') %>%
  tabyl(Year, Trip_Type) %>%
  adorn_totals(c("row", "col"))


```

## Sample Sizes by Mode

```{r}
#| echo: false
#| eval: true

#unique(lengths$State_Landed)



paste0("Table: Number of discarded fish lengths from each mode")
lengths %>%
  filter(Disposition == 'Discard') %>%
  tabyl(Year, Fishing_Mode) %>%
  adorn_totals(c("row", "col"))



```

## Sample Sizes by Data Provider and Mode - Discards

```{r}
#| echo: false
#| eval: true

#unique(lengths$State_Landed)



paste0("Table: Number of discarded fish lengths from each trip type.")
lengths %>%
  filter(Disposition == 'Discard') %>%
  tabyl(Year, Data_Provider, Fishing_Mode) %>%
  adorn_totals(c("row", "col"))



```

## Sample Sizes by Potential Weighting Subregion - Charter

```{r}
#| echo: false
#| eval: false



paste0("Table: Number of discarded fish by weighting subregion.")
lengths %>%
  filter(Fishing_Mode == 'CB') %>%
  filter(Disposition == 'Discard') %>%
  tabyl(Year, Sub_Reg) %>%
  adorn_totals(c("row", "col"))


```

## Length Distribution by State - Discards

### Years Aggregated

```{r}
#| echo: false
#| fig-cap: "Figure. Comparison of discard length distribution by state."
#| warning: false
#| eval: false


discards <- lengths %>%
  filter(Disposition == 'Discard')


max(discards$Observed_FL_mm, na.rm=TRUE)

#discards$State_Landed <- factor(discards$State_Landed,
                               levels=c('MS', 'AL', 'FL'))


state.dis.obs <- discards %>% 
  left_join(lengths %>% group_by(State_Landed) %>% summarise(N=n()))%>%
  mutate(Label=paste0(State_Landed,' (n = ',N,')'))







ggplot(state.dis.obs, aes(x=Final_Length_mm)) +
       geom_density(linewidth=1) +
       xlab(paste0(length.type)) +
       ylab("Density") +
       labs(title = "")+ 
       #scale_color_manual(name = 'State', 
         #    values = c('deeppink', 'darkblue', 'darkorange')) +
  scale_x_continuous(breaks=seq(100,900,100), limits=c(0,950), expand=c(0,0)) +
    #facet_wrap(~as.factor(Year), ncol=8, scales = 'freey') +
  theme_minimal() +
  theme(strip.background = element_rect(fill="white", size=1)) +
  theme(strip.text = element_text(size=9)) +
  theme(plot.title = element_text(hjust=0.5)) +
  theme(legend.position = 'bottom') +
  theme(legend.title=element_blank()) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  theme(axis.text=element_text(size=8),
        text=element_text(size=9)) 
```

### Annual

```{r}
#| echo: false
#| fig-cap: "Figure. Comparison of annual discard length distribution by state. Note differences in y-axis scale"
#| fig-width: 11
#| fig-height: 8.5
#| warning: false
#| label: 'state_annual_discards'
#| eval: false



state.nobs <- discards %>%
  group_by(Year, State_Landed) %>%
  summarise(N = n())

#ms.nobs <-state.nobs %>%
#  filter(State_Landed == 'MS')

#al.nobs <-state.nobs %>%
#  filter(State_Landed == 'AL')

fl.nobs <- state.nobs %>%
  filter(State_Landed == 'FL')



ggplot(discards, aes(x=Final_Length_mm, colour=State_Landed)) +
       geom_density(linewidth=1) +
       xlab(paste0(length.type)) +
       ylab("Density") +
       labs(title = "")+ 
       scale_color_manual(name = 'State', 
              values = c('deeppink', 'darkorange', 'darkblue')) +
  scale_x_continuous(breaks=seq(0,900,100), limits=c(0,850), expand=c(0,0)) +
  #geom_text(aes(label=paste0('n=', N)), x=Inf, y=Inf, vjust=1.1, hjust=1.2, data=ms.nobs, colour ='deeppink', size=3) +
   # geom_text(aes(label=paste0('n=', N)), x=Inf, y=Inf, vjust=2.3, hjust=1.2, data=al.nobs, colour ='darkorange', size=3) +
   geom_text(aes(label=paste0('n=', N)), x=Inf, y=Inf, vjust=3.5, hjust=1.2, data=fl.nobs, colour ='darkblue', size=3) +
  facet_wrap(~as.factor(Year), ncol=5, scale='free_y') +
  theme_minimal() +
  theme(strip.background = element_rect(fill="white", size=1)) +
  theme(strip.text = element_text(size=9)) +
  theme(plot.title = element_text(hjust=0.5)) +
  theme(legend.position = 'bottom') +
  theme(legend.title=element_blank()) +
  theme(axis.text=element_text(size=7),
        text=element_text(size=7)) 
```

## Length Distribution by FL Reg

```{r}
#| echo: false
#| fig-cap: "Figure. Comparison of annual discard length distribution by FL region. Note differences in y-axis scale"
#| fig-width: 11
#| fig-height: 8.5
#| warning: false
#| label: 'fl_reg_annual_discards'
#| eval: false




discards.fl <- discards %>%
  filter(State_Landed == 'FL')

unique(discards.fl$hb_reg)


fl.nobs <- discards.fl %>%
  group_by(Year, hb_reg) %>%
  summarise(N = n())

nwfl.nobs <-fl.nobs %>%
  filter(hb_reg == 'NWFL')

swfl.nobs <-fl.nobs %>%
  filter(hb_reg == 'SWFL')




##* *Order regions geographically*
discards.fl$hb_reg <- factor(discards.fl$hb_reg,
                               levels=c('NWFL', 'SWFL'))




ggplot(discards.fl, aes(x=Final_Length_mm, colour=hb_reg)) +
       geom_density(linewidth=1) +
       xlab(paste0(length.type)) +
       ylab("Density") +
       labs(title = "")+ 
       scale_color_manual(name = 'hb_reg', 
              values = c('deeppink', 'purple4', 'darkorange', 'darkblue')) +
  scale_x_continuous(breaks=seq(100,600,100), limits=c(0,650), expand=c(0,0)) +
  geom_text(aes(label=paste0('n=', N)), x=Inf, y=Inf, vjust=1.1, hjust=1.2, data=nwfl.nobs, colour ='deeppink', size=3) +
   geom_text(aes(label=paste0('n=', N)), x=Inf, y=Inf, vjust=2.3, hjust=1.2, data=swfl.nobs, colour ='purple4', size=3) +
   #geom_text(aes(label=paste0('n=', N)), x=Inf, y=Inf, vjust=3.5, hjust=1.2, data=keys.nobs, colour ='darkblue', size=3) +
  facet_wrap(~as.factor(Year), ncol=5, scale='free_y') +
  theme_minimal() +
  theme(strip.background = element_rect(fill="white", size=1)) +
  theme(strip.text = element_text(size=9)) +
  theme(plot.title = element_text(hjust=0.5)) +
  theme(legend.position = 'bottom') +
  theme(legend.title=element_blank()) +
  theme(axis.text=element_text(size=7),
        text=element_text(size=7)) 
```

## Length Distribution by Potential Weighting Subregion

```{r}
#| echo: false
#| fig-cap: "Figure. Comparison of annual discard length distribution by potential subweighting region. Note differences in y-axis scale"
#| fig-width: 11
#| fig-height: 8.5
#| warning: false
#| label: 'subweighting_reg_annual_discards_charter'
#| eval: false





subreg.nobs <- discards %>%
  group_by(Year, Sub_Reg) %>%
  summarise(N = n())

n.nobs <- subreg.nobs %>%
  filter(Sub_Reg == 'North')

s.nobs <- subreg.nobs %>%
  filter(Sub_Reg == 'South')




##* *Order regions geographically*
#discards.fl$hb_reg <- factor(discards.fl$hb_reg,
#                               levels=c('NWFL', 'SWFL'))


cb.discards <- discards %>%
  filter(Fishing_Mode == 'CB')


ggplot(cb.discards, aes(x=Final_Length_mm, colour=Sub_Reg)) +
       geom_density(linewidth=1) +
       xlab(paste0(length.type)) +
       ylab("Density") +
       labs(title = "")+ 
       scale_color_manual(name = 'hb_reg', 
              values = c('deeppink', 'purple4', 'darkorange', 'darkblue')) +
  scale_x_continuous(breaks=seq(100,600,100), limits=c(0,650), expand=c(0,0)) +
  geom_text(aes(label=paste0('n=', N)), x=Inf, y=Inf, vjust=1.1, hjust=1.2, data=n.nobs, colour ='deeppink', size=3) +
   geom_text(aes(label=paste0('n=', N)), x=Inf, y=Inf, vjust=2.3, hjust=1.2, data=s.nobs, colour ='purple4', size=3) +
   #geom_text(aes(label=paste0('n=', N)), x=Inf, y=Inf, vjust=3.5, hjust=1.2, data=keys.nobs, colour ='darkblue', size=3) +
  facet_wrap(~as.factor(Year), ncol=5, scale='free_y') +
  theme_minimal() +
  theme(strip.background = element_rect(fill="white", size=1)) +
  theme(strip.text = element_text(size=9)) +
  theme(plot.title = element_text(hjust=0.5)) +
  theme(legend.position = 'bottom') +
  theme(legend.title=element_blank()) +
  theme(axis.text=element_text(size=7),
        text=element_text(size=7)) 

```

## Length Distribution by Trip Type

```{r}
#| echo: false
#| fig-cap: "Figure. Comparison of annual discard length distribution by FL region. Note differences in y-axis scale."
#| fig-width: 11
#| fig-height: 8.5
#| warning: false
#| label: 'trip_type_discards'
#| eval: false




type.nobs <- discards %>%
  group_by(Year, trip_type) %>%
  summarise(N = n())

h.nobs <-type.nobs %>%
  filter(trip_type == 'H')

q.nobs <-type.nobs %>%
  filter(trip_type == 'Q')

f.nobs <-type.nobs %>%
  filter(trip_type == 'F')

m.nobs <-type.nobs %>%
  filter(trip_type == 'M')




ggplot(discards, aes(x=Final_Length_mm, colour=trip_type)) +
       geom_density(linewidth=1) +
       xlab("Max TL (mm)") +
       ylab("Density") +
       labs(title = "")+ 
       scale_color_manual(name = 'trip_type', 
              values = c('deeppink', 'darkorange', 'darkblue','purple')) +
  scale_x_continuous(breaks=seq(250,1300,250), limits=c(50,1400), expand=c(0,0)) +
  geom_text(aes(label=paste0('n=', N)), x=Inf, y=Inf, vjust=1.1, hjust=1.2, data=h.nobs, colour ='deeppink', size=3) +
   geom_text(aes(label=paste0('n=', N)), x=Inf, y=Inf, vjust=2.3, hjust=1.2, data=q.nobs, colour ='darkorange', size=3) +
   geom_text(aes(label=paste0('n=', N)), x=Inf, y=Inf, vjust=3.5, hjust=1.2, data=f.nobs, colour ='darkblue', size=3) +
  geom_text(aes(label=paste0('n=', N)), x=Inf, y=Inf, vjust=4.7, hjust=1.2, data=m.nobs, colour ='purple', size=3) +
  facet_wrap(~as.factor(Year), ncol=5, scale='free_y') +
  theme_minimal() +
  theme(strip.background = element_rect(fill="white", size=1)) +
  theme(strip.text = element_text(size=9)) +
  theme(plot.title = element_text(hjust=0.5)) +
  theme(legend.position = 'bottom') +
  theme(legend.title=element_blank()) +
  theme(axis.text=element_text(size=7),
        text=element_text(size=7)) 
```

## Length Distribution by Mode

```{r}
#| echo: false
#| fig-cap: "Figure. Comparison of annual discard length distribution by FL region. Note differences in y-axis scale"
#| fig-width: 11
#| fig-height: 8.5
#| warning: false
#| label: 'mode_discards'
#| eval: false




#type.nobs <- discards %>%
#  group_by(Year, trip_type) %>%
#  summarise(N = n())

#h.nobs <-type.nobs %>%
#  filter(trip_type == 'H')

#q.nobs <-type.nobs %>%
#  filter(trip_type == 'Q')

#f.nobs <-type.nobs %>%
#  filter(trip_type == 'F')

#m.nobs <-type.nobs %>%
#  filter(trip_type == 'M')




ggplot(discards, aes(x=Final_Length_mm, colour=Fishing_Mode)) +
       geom_density(linewidth=1) +
       xlab(paste0(length.type)) +
       ylab("Density") +
       labs(title = "")+ 
       scale_color_manual(name = 'trip_type', 
              values = c('deeppink', 'darkorange', 'darkblue','purple')) +
  scale_x_continuous(breaks=seq(100,900,100), limits=c(0,950), expand=c(0,0)) +
#  geom_text(aes(label=paste0('n=', N)), x=Inf, y=Inf, vjust=1.1, hjust=1.2, data=h.nobs, colour ='deeppink', size=3) +
#   geom_text(aes(label=paste0('n=', N)), x=Inf, y=Inf, vjust=2.3, hjust=1.2, data=q.nobs, colour ='darkorange', size=3) +
#   geom_text(aes(label=paste0('n=', N)), x=Inf, y=Inf, vjust=3.5, hjust=1.2, data=f.nobs, colour ='darkblue', size=3) +
#  geom_text(aes(label=paste0('n=', N)), x=Inf, y=Inf, vjust=4.7, hjust=1.2, data=m.nobs, colour ='purple', size=3) +
  facet_wrap(~as.factor(Year), ncol=5, scale='free_y') +
  theme_minimal() +
  theme(strip.background = element_rect(fill="white", size=1)) +
  theme(strip.text = element_text(size=9)) +
  theme(plot.title = element_text(hjust=0.5)) +
  theme(legend.position = 'bottom') +
  theme(legend.title=element_blank()) +
  theme(axis.text=element_text(size=7),
        text=element_text(size=7)) 
```

# Harvested vs Discarded Lengths

## Headboat

```{r}
#| output: false
#| echo: false
#| warning: false
#| eval: true




##* *Summarize number of harvested and discarded lengths by 1 inch length bin*


lengths2 <- lengths %>%
  #filter(Fishing_Mode == 'HB') %>%
  group_by(Year, Disposition, bin_1in) %>%
  summarise(nfish = n())



##* *Summarize by disposition type to add to figures*
disposition.n <- lengths %>%
  #filter(Fishing_Mode == 'HB') %>%
  group_by(Year, Disposition) %>%
  summarise(nfish = n())


harvest.n <- disposition.n %>%
  filter(Disposition == 'Harvest')

harvest <- harvest.n[,]

discard.n <- disposition.n %>%
  filter(Disposition == 'Discard')


```

```{r}
#| echo: false
#| fig-cap: "Figure. Comparison of harvested and discarded length distributions."
#| fig-align: 'left'
#| warning: false
#| label: 'annual_harvest_vs_discard_hb'
#| eval: false



ggplot(dat=lengths2, aes(x=bin_1in, y=nfish, fill=Disposition)) +
    geom_bar(stat='identity')+
    xlab("FL (in)") +
    ylab("n") +
    #geom_vline(xintercept=20, linetype='dashed', color='black') +
    #labs(title = "Nominal Length Compositions - GR") +
    #scale_x_continuous(breaks=seq(10, 50, by=10), limits=c(0, 50), expand=c(0,0)) +
  scale_fill_manual(name = 'Disposition', 
             values = c('red', 'darkblue')) +
    facet_wrap(~as.factor(Year), ncol=5, scales='free_y') +
    geom_text(aes(label=paste0('n=', nfish)), x=Inf, y=Inf, vjust=1.1, hjust=1.2, data=discard.n, colour ='red', size=3) +
    geom_text(aes(label=paste0('n=', nfish)), x=Inf, y=Inf, vjust=2.1, hjust=1.2, data=harvest.n, colour ='darkblue', size=3) +
    theme_minimal() +
    theme(strip.background = element_rect(fill="white", size=1)) +
    theme(strip.text = element_text(size=8)) +
    theme(plot.title = element_text(hjust=0.5)) +
    theme(legend.position = 'bottom') +
    #theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
    theme(axis.text=element_text(size=8),
          text=element_text(size=8)) 

```

## Charter

```{r}
#| output: false
#| echo: false
#| warning: false
#| eval: false




##* *Summarize number of harvested and discarded lengths by 1 inch length bin*


lengths2 <- lengths %>%
  filter(Fishing_Mode == 'CB') %>%
  group_by(Year, Disposition, bin_1in) %>%
  summarise(nfish = n())



##* *Summarize by disposition type to add to figures*
disposition.n <- lengths %>%
  filter(Fishing_Mode == 'CB') %>%
  group_by(Year, Disposition) %>%
  summarise(nfish = n())


harvest.n <- disposition.n %>%
  filter(Disposition == 'Harvest')

harvest <- harvest.n[,]

discard.n <- disposition.n %>%
  filter(Disposition == 'Discard')



```

```{r}
#| echo: false
#| fig-cap: "Figure. Comparison of harvested and discarded length distributions."
#| fig-align: 'left'
#| warning: false
#| label: 'annual_harvest_vs_discard_cb'
#| eval: false



ggplot(dat=lengths2, aes(x=bin_1in, y=nfish, fill=Disposition)) +
    geom_bar(stat='identity')+
    xlab("FL (in)") +
    ylab("n") +
    #geom_vline(xintercept=20, linetype='dashed', color='black') +
    #labs(title = "Nominal Length Compositions - GR") +
    #scale_x_continuous(breaks=seq(10, 50, by=10), limits=c(0, 50), expand=c(0,0)) +
  scale_fill_manual(name = 'Disposition', 
             values = c('red', 'darkblue')) +
    facet_wrap(~as.factor(Year), ncol=5, scales='free_y') +
    geom_text(aes(label=paste0('n=', nfish)), x=Inf, y=Inf, vjust=1.1, hjust=1.2, data=discard.n, colour ='red', size=3) +
    geom_text(aes(label=paste0('n=', nfish)), x=Inf, y=Inf, vjust=2.1, hjust=1.2, data=harvest.n, colour ='darkblue', size=3) +
    theme_minimal() +
    theme(strip.background = element_rect(fill="white", size=1)) +
    theme(strip.text = element_text(size=8)) +
    theme(plot.title = element_text(hjust=0.5)) +
    theme(legend.position = 'bottom') +
    #theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
    theme(axis.text=element_text(size=8),
          text=element_text(size=8)) 


```

```{r}
#| echo: false
#| eval: false

##* *Export the length data that has bins assigned*
##* *No samples removed from data*

out.name <- paste0("Discard_LenComps/S99_KM_obs_len_final_", Sys.Date(), ".xlsx")


write.xlsx(lengths, 
           file = out.name, 
           sheetName = "KM", 
           colNames = TRUE, 
           rowNames = FALSE, 
           append = FALSE,
           overwrite=TRUE) 


```

# Nominal Discard Length Comps

## 2 cm bins

```{r}
#| echo: false
#| output: false
#| eval: true


##* *Aggregated comps - 2 cm*


##* *Nominal Discard Length Comps - Rec - MS-FL*
nom.l.wide.rec <- readxl::read_xlsx(paste0(dir,'/DiscardLenComps/SBS_rec_dlfd_agg_singletab_0423_20260611.xlsx'), sheet = '2cm')

##* *Remove years with n_fish < 30*
#nom.l.wide <- subset(nom.l.wide, n_fish >=30)
  
  
##* *Reorganize the data so the lengths are rows instead of columns*
nom.l.rec.g <- nom.l.wide.rec %>%
  pivot_longer(cols = !c(Fleet, n_fish, n_trip, Time_Block),
               names_to = "bin",
               values_to = "Len_Comp")

nom.l.rec.g$Comp <- "Nominal"


nom.l.rec.g$bin <- as.numeric(nom.l.rec.g$bin)


nom.l.rec.g <- nom.l.rec.g %>%
  filter(Time_Block != 'pooled')


```

```{r}
#| echo: false
#| fig-cap: "Figure. Annual nominal discard length composition. Note differences in y-axis scale"
#| fig-width: 11
#| fig-height: 8.5
#| warning: false
#| label: 'nom_discard_2cm'
#| eval: true



disc.nobs <- nom.l.wide.rec[,c(1,2,3,4)] 

t1.nobs <-disc.nobs %>%
  filter(Time_Block == '2004to2007')

t2.nobs <-disc.nobs %>%
  filter(Time_Block == '2008to2023')

pooled.nobs <-disc.nobs %>%
  filter(Time_Block == 'pooled')


length.type <- 'FL (cm)'

ggplot(dat=nom.l.rec.g, aes(x=bin, y=Len_Comp, colour=Time_Block)) +
    geom_line(linewidth=1) +
         xlab("Bin, FL (cm)") +
         ylab("Proportion of Annual Sample") +
         labs(title = "") +
         scale_color_manual(name = 'Comp', 
                 values = c('magenta', 'darkblue', 'deeppink')) +
  #scale_x_continuous(breaks=seq(20, 100, by=20), limits=c(0, 120), expand=c(0,0)) +
  geom_text(aes(label=paste0('n=', n_fish)), x=Inf, y=Inf, vjust=1.1, hjust=1.2, data=t1.nobs, colour ='magenta2', size=3) +
  geom_text(aes(label=paste0('n=', n_fish)), x=Inf, y=Inf, vjust=2.1, hjust=1.2, data=t2.nobs, colour ='darkblue', size=3) +
  #geom_text(aes(label=paste0('n=', n_fish)), x=Inf, y=Inf, vjust=3.1, hjust=1.2, data=pooled.nobs, colour ='deeppink', size=3) +
  #facet_wrap(~as.factor(Year), ncol=5, scales='free_y') +
  theme_minimal() +
  theme(strip.background = element_rect(fill="white", size=1)) +
  theme(strip.text = element_text(size=9)) +
  theme(plot.title = element_text(hjust=0.5)) +
  theme(legend.position = 'bottom') +
  theme(axis.text=element_text(size=8),
        text=element_text(size=9)) 




```

# 

# 

```{r}
#| echo: false
#| fig-cap: "Figure. Annual nominal discard length composition. Note differences in y-axis scale"
#| warning: false
#| label: 'nom_discard_2cm_hist'
#| eval: true



disc.nobs <- nom.l.wide.rec[,c(1,2,3,4)] 

t1.nobs <-disc.nobs %>%
  filter(Time_Block == '2004to2007')

t2.nobs <-disc.nobs %>%
  filter(Time_Block == '2008to2023')

pooled.nobs <-disc.nobs %>%
  filter(Time_Block == 'pooled')


length.type <- 'FL (cm)'


lbin.type <- 'bin, (FL)'

ggplot(dat=nom.l.rec.g, aes(x=bin, y=Len_Comp, fill=Time_Block)) +
    geom_histogram(position = "identity", stat="identity", alpha=0.6) +
         xlab(paste0(lbin.type)) +
         ylab("Proportion of Annual Sample") +
         labs(title = "Nominal Discard Length - 2 cm bins") +
         scale_fill_manual(name = 'Time Block', 
                 values = c('magenta', 'navy')) +
scale_x_continuous(breaks=seq(0, 130, by=10), limits=c(0, 132), expand=c(0,0)) +
  geom_text(aes(label=paste0('n=', n_fish)), x=Inf, y=Inf, vjust=1.1, hjust=1.2, data=t1.nobs, colour ='magenta', size=3) +
  geom_text(aes(label=paste0('n=', n_fish)), x=Inf, y=Inf, vjust=2.1, hjust=1.2, data=t2.nobs, colour ='navy', size=3) +
  #geom_text(aes(label=paste0('n=', n_fish)), x=Inf, y=Inf, vjust=3.1, hjust=1.2, data=prsh.nobs, colour ='blue2', size=3) +
  #facet_wrap(~as.factor(Year), ncol=5, scales='free_y') +
  theme_minimal() +
  #theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust=1)) +
  theme(strip.background = element_rect(fill="white", size=1)) +
  theme(strip.text = element_text(size=9)) +
  theme(plot.title = element_text(hjust=0.5)) +
  theme(legend.position = 'bottom') +
  theme(axis.text=element_text(size=8),
        text=element_text(size=8)) 




```

## 5 cm

```{r}
#| echo: false
#| output: false
#| eval: true


##* *Aggregated comps - 5 cm*


##* *Nominal Discard Length Comps - Rec - MS-FL*
nom.l.wide.rec <- readxl::read_xlsx(paste0(dir,'/DiscardLenComps/SBS_rec_dlfd_agg_singletab_0423_20260611.xlsx'), sheet = '5cm')

##* *Remove years with n_fish < 30*
#nom.l.wide <- subset(nom.l.wide, n_fish >=30)
  
  
##* *Reorganize the data so the lengths are rows instead of columns*
nom.l.rec.g <- nom.l.wide.rec %>%
  pivot_longer(cols = !c(Fleet, n_fish, n_trip, Time_Block),
               names_to = "bin",
               values_to = "Len_Comp")

nom.l.rec.g$Comp <- "Nominal"


nom.l.rec.g$bin <- as.numeric(nom.l.rec.g$bin)


nom.l.rec.g <- nom.l.rec.g %>%
  filter(Time_Block != 'pooled')



```

```{r}
#| echo: false
#| fig-cap: "Figure. Annual nominal discard length composition. Note differences in y-axis scale"
#| warning: false
#| label: 'nom_discard_5cm_hist'
#| eval: true



disc.nobs <- nom.l.wide.rec[,c(1,2,3,4)] 

t1.nobs <-disc.nobs %>%
  filter(Time_Block == '2004to2007')

t2.nobs <-disc.nobs %>%
  filter(Time_Block == '2008to2023')

pooled.nobs <-disc.nobs %>%
  filter(Time_Block == 'pooled')


length.type <- 'FL (cm)'


lbin.type <- 'bin, (FL)'

ggplot(dat=nom.l.rec.g, aes(x=bin, y=Len_Comp, fill=Time_Block)) +
    geom_histogram(position = "identity", stat="identity", alpha=0.6) +
         xlab(paste0(lbin.type)) +
         ylab("Proportion of Annual Sample") +
         labs(title = "Nominal Discard Length - 5 cm bins") +
         scale_fill_manual(name = 'Time Block', 
                 values = c('magenta', 'navy')) +
scale_x_continuous(breaks=seq(0, 130, by=10), limits=c(0, 132), expand=c(0,0)) +
  geom_text(aes(label=paste0('n=', n_fish)), x=Inf, y=Inf, vjust=1.1, hjust=1.2, data=t1.nobs, colour ='magenta', size=3) +
  geom_text(aes(label=paste0('n=', n_fish)), x=Inf, y=Inf, vjust=2.1, hjust=1.2, data=t2.nobs, colour ='navy', size=3) +
  #geom_text(aes(label=paste0('n=', n_fish)), x=Inf, y=Inf, vjust=3.1, hjust=1.2, data=prsh.nobs, colour ='blue2', size=3) +
  #facet_wrap(~as.factor(Year), ncol=5, scales='free_y') +
  theme_minimal() +
  #theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust=1)) +
  theme(strip.background = element_rect(fill="white", size=1)) +
  theme(strip.text = element_text(size=9)) +
  theme(plot.title = element_text(hjust=0.5)) +
  theme(legend.position = 'bottom') +
  theme(axis.text=element_text(size=8),
        text=element_text(size=8)) 





```

## 10 cm

```{r}
#| echo: false
#| output: false
#| eval: true


##* *Aggregated comps - 10 cm*


##* *Nominal Discard Length Comps - Rec - MS-FL*
nom.l.wide.rec <- readxl::read_xlsx(paste0(dir,'/DiscardLenComps/SBS_rec_dlfd_agg_singletab_0423_20260611.xlsx'), sheet = '10cm')

##* *Remove years with n_fish < 30*
#nom.l.wide <- subset(nom.l.wide, n_fish >=30)
  
  
##* *Reorganize the data so the lengths are rows instead of columns*
nom.l.rec.g <- nom.l.wide.rec %>%
  pivot_longer(cols = !c(Fleet, n_fish, n_trip, Time_Block),
               names_to = "bin",
               values_to = "Len_Comp")

nom.l.rec.g$Comp <- "Nominal"


nom.l.rec.g$bin <- as.numeric(nom.l.rec.g$bin)


nom.l.rec.g <- nom.l.rec.g %>%
  filter(Time_Block != 'pooled')


```

```{r}
#| echo: false
#| fig-cap: "Figure. Annual nominal discard length composition. Note differences in y-axis scale"
#| warning: false
#| label: 'nom_discard_10cm_hist'
#| eval: true



disc.nobs <- nom.l.wide.rec[,c(1,2,3,4)] 

t1.nobs <-disc.nobs %>%
  filter(Time_Block == '2004to2007')

t2.nobs <-disc.nobs %>%
  filter(Time_Block == '2008to2023')

pooled.nobs <-disc.nobs %>%
  filter(Time_Block == 'pooled')


length.type <- 'FL (cm)'


lbin.type <- 'bin, (FL)'

ggplot(dat=nom.l.rec.g, aes(x=bin, y=Len_Comp, fill=Time_Block)) +
    geom_histogram(position = "identity", stat="identity", alpha=0.6) +
         xlab(paste0(lbin.type)) +
         ylab("Proportion of Annual Sample") +
         labs(title = "Nominal Discard Length - 10 cm bins") +
         scale_fill_manual(name = 'Time Block', 
                 values = c('magenta', 'navy')) +
scale_x_continuous(breaks=seq(0, 130, by=10), limits=c(0, 132), expand=c(0,0)) +
  geom_text(aes(label=paste0('n=', n_fish)), x=Inf, y=Inf, vjust=1.1, hjust=1.2, data=t1.nobs, colour ='magenta', size=3) +
  geom_text(aes(label=paste0('n=', n_fish)), x=Inf, y=Inf, vjust=2.1, hjust=1.2, data=t2.nobs, colour ='navy', size=3) +
  #geom_text(aes(label=paste0('n=', n_fish)), x=Inf, y=Inf, vjust=3.1, hjust=1.2, data=prsh.nobs, colour ='blue2', size=3) +
  #facet_wrap(~as.factor(Year), ncol=5, scales='free_y') +
  theme_minimal() +
  #theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust=1)) +
  theme(strip.background = element_rect(fill="white", size=1)) +
  theme(strip.text = element_text(size=9)) +
  theme(plot.title = element_text(hjust=0.5)) +
  theme(legend.position = 'bottom') +
  theme(axis.text=element_text(size=8),
        text=element_text(size=8)) 




```

# Nominal and Weighted Comparisons

## FL Only

```{r}
#| echo: false
#| output: false
#| eval: false


##* *FL nominal and weighted length comps*
FL.wgt.wide2 <- readxl::read_xlsx(paste0(dir,'/DiscardLenComps/RS_rec_dlfdw_FL_hb_0524_20250612.xlsx'), sheet = 'comp2')

##* *Remove years with n_fish < 30*
#nom.l.wide <- subset(nom.l.wide, n_fish >=30)
  
  
  
##* *Reorganize the data so the lengths are rows instead of columns*
FL.comp <- FL.wgt.wide2 %>%
  pivot_longer(cols = !c(Year, n_fish, n_trip, Type),
               names_to = "bin",
               values_to = "Len_Comp")



FL.comp$bin <- as.numeric(FL.comp$bin)






```

```{r}
#| echo: false
#| fig-cap: "Figure. Comparison of weighting methods and nominal for FL headboat discard length comps. Note differences in y-axis scale"
#| fig-width: 11
#| fig-height: 8.5
#| warning: false
#| label: 'FL_weighted_comparisons_nom'
#| eval: false




ggplot(dat=FL.comp, aes(x=bin, y=Len_Comp, colour=Type)) +
    geom_line(linewidth=1) +
         xlab("Bin, paste0(length.type)") +
         ylab("Proportion of Annual Sample") +
         labs(title = "") +
         scale_color_manual(name = 'Comp', 
                 values = c('green', 'darkblue', 'deeppink')) +
  scale_x_continuous(breaks=seq(20, 100, by=20), limits=c(0, 120), expand=c(0,0)) +
  facet_wrap(~as.factor(Year), ncol=5, scales='free_y') +
  theme_minimal() +
  theme(strip.background = element_rect(fill="white", size=1)) +
  theme(strip.text = element_text(size=9)) +
  theme(plot.title = element_text(hjust=0.5)) +
  theme(legend.position = 'bottom') +
  theme(axis.text=element_text(size=8),
        text=element_text(size=9)) 


```

# 

## All Data

```{r}
#| echo: false
#| output: false
#| eval: false


##* *Single Gulf Stock*


##* *Weighted Discard Length Comps - Rec - MS-FL*
wgt.l.wide.rec <- readxl::read_xlsx(paste0(dir,'/DiscardLenComps/GTF_rec_dlfdw_0524_20260121.xlsx'), sheet = 'GR')

##* *Remove years with n_fish < 30*
#nom.l.wide <- subset(nom.l.wide, n_fish >=30)
  
##* *Remove years with n_fish = 0*
wgt.l.wide.rec <- wgt.l.wide.rec %>%
  filter(n_fish >=1)
  
##* *Reorganize the data so the lengths are rows instead of columns*
wgt.l.rec.g <-wgt.l.wide.rec %>%
  pivot_longer(cols = !c(Year, Fleet, n_fish, n_trip),
               names_to = "bin",
               values_to = "Len_Comp")

wgt.l.rec.g$Comp <- "Weighted"


wgt.l.rec.g$bin <- as.numeric(wgt.l.rec.g$bin)




##* *Combine nominal and weighted comps*
disc.comps <- rbind(nom.l.rec.g, wgt.l.rec.g)




nom.nobs <-disc.comps %>%
  filter(Comp == 'Nominal')

wgt.nobs <-disc.comps %>%
  filter(Comp == 'Weighted')



```

```{r}
#| echo: false
#| fig-cap: "Figure. Comparison of nominal and weighted discard length comps. Note differences in y-axis scale"
#| fig-width: 11
#| fig-height: 8.5
#| warning: false
#| label: 'nom_vs_weighted'
#| eval: false




ggplot(dat=disc.comps, aes(x=bin, y=Len_Comp, colour=Comp)) +
    geom_line(linewidth=1) +
         xlab("Bin (Fork Length cm)") +
         ylab("Proportion of Annual Sample") +
         labs(title = "") +
         scale_color_manual(name = 'Comp', 
                 values = c('magenta2', 'darkblue', 'deeppink')) +
  scale_x_continuous(breaks=seq(0, 60, by=10), limits=c(0, 62), expand=c(0,0)) +
  geom_text(aes(label=paste0('n=', n_fish)), x=Inf, y=Inf, vjust=1.1, hjust=1.2, data=nom.nobs, colour ='magenta2', size=3) +
  geom_text(aes(label=paste0('n=', n_fish)), x=Inf, y=Inf, vjust=2.1, hjust=1.2, data=wgt.nobs, colour ='darkblue', size=3) +
  facet_wrap(~as.factor(Year), ncol=5, scales='free_y') +
  theme_minimal() +
  theme(strip.background = element_rect(fill="white", size=1)) +
  theme(strip.text = element_text(size=9)) +
  theme(plot.title = element_text(hjust=0.5)) +
  theme(legend.position = 'bottom') +
  theme(axis.text=element_text(size=8),
        text=element_text(size=9)) 



```

## Sample/Trip Sizes Using Ellie's New File (3/11)

```{r}
#| echo: false
#| output: false
#| eval: false


##* *Define directory*
dir <- getwd()

##* *Prevents R from converting any long numbers (ID codes) to scientific notation*
options(scipen=999)


##* *Import data*
fl.lengths <- readxl::read_xlsx(paste0(dir,'/Discard_LenComps/KingMackerel_observerlengs_raw_FWRIv3.xlsx'), sheet = 'Headboat Raw Lengths', guess_max = 500000)



##* *Assign fishing year*
fl.lengths$Month2 <- as.numeric(fl.lengths$Month)

fl.lengths$Year2 <- as.numeric(fl.lengths$YEAR)



fl.lengths <- fl.lengths %>%
  mutate(Year = case_when(
    Month2 %in% c(7,8,9,10,11,12) ~ Year2,
    Month2 %in% c(1,2,3,4,5,6)    ~ Year2-1))



```

```{r}
#| echo: false
#| output: false
#| eval: false


hb.trips <- fl.lengths %>%
  filter(catchtype == 'D') %>%
  group_by(Year) %>%
  summarise(n_fish = n(),
         n_trip = length(unique(Sampling_Unit_ID)))


```

```{r}
#| echo: false
#| output: false
#| eval: false


##* *Define directory*
dir <- getwd()

##* *Prevents R from converting any long numbers (ID codes) to scientific notation*
options(scipen=999)


##* *Import data*
fl.lengths <- readxl::read_xlsx(paste0(dir,'/Discard_LenComps/KingMackerel_observerlengs_raw_FWRIv3.xlsx'), sheet = 'Charter Raw Lengths', guess_max = 500000)



##* *Assign fishing year*
fl.lengths$Month2 <- as.numeric(fl.lengths$Month)

fl.lengths$Year2 <- as.numeric(fl.lengths$YEAR)



fl.lengths <- fl.lengths %>%
  mutate(Year = case_when(
    Month2 %in% c(7,8,9,10,11,12) ~ Year2,
    Month2 %in% c(1,2,3,4,5,6)    ~ Year2-1))



```

```{r}
#| echo: false
#| output: false
#| eval: false


cb.trips <- fl.lengths %>%
  filter(catchtype == 'D') %>%
  group_by(Year) %>%
  summarise(n_fish = n(),
         n_trip = length(unique(Sampling_Unit_ID)))


```
